---
title: "Wrangle rows and columns"
format: html
---

# Instructions

This week, you learned several more functions from the dplyr package (pronounced DEE-ply-er). These functions help you work with the rows and columns in your data set.

You'll use these skills to investigate the effect on solar potential of angling a solar panel to face the sun.

To learn more about the variables contained in the data set, refer to the data dictionary: https://rsacdn.link/milestones/internal/v2/tidyverse/solar-power/assets/illuminance_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false

# Load your packages here

```

## Create and select columns

In this milestone, you'll create two new variables, one that measures the direct solar illuminance received by a panel laid flat on the ground, and another that measures the difference in illuminance received by a flat panel vs a panel that is angled towards the sun. You'll then use these variables to assess the benefits of angling a solar panel towards the sun.

## Recreation

### Part 1 - Import and clean

Before you begin, you will need to import your data set. Use the code chunk below to read the data from the data file `illuminance.csv`, which is stored in the `data/` folder in your working directory. Be sure to save the data to an object named `illuminance`.

```{r}
#| label: recreation-import

```

You now have the tools to rename your columns with transparent, human readable names. Let's do that before we move on. To get started, run the code below to see a table.

```{r}
#| label: recreate-this-a
#| message: false
solution_a <- readr::read_csv("data/milestone03a.csv")
solution_a
```

Your task is to use what you've learned about the dplyr package to transform `illuminance` into this table. You will need to replace the column names on the right with the new names on the left.

| New name          | Old name             |
|-------------------|----------------------|
| global_lux        | global_illuminance   |
| diffuse_lux       | diffuse_illuminance  |
| direct_lux        | direct_illuminance   |
| pressure          | ps                   |
| temp              | t2m                  |
| temp_min          | t2m_min              |
| temp_max          | t2m_max              |
| dew_point         | t2mdew               |
| snow              | snodp                |
| wind_speed        | ws2m                 |
| wind_direction    | wd2m                 |
| uv_index          | allsky_sfc_uv_index  |
| albedo            | allsky_srf_alb       |
| optical_depth_55  | aod_55               |
| optical_depth_84  | aod_84               |
| clearness         | clrsky_kt            |

Work in the code chunk below. Save the result as `solar`. We will use `solar` in place of `illuminance` going forward.

```{r}
#| label: recreation-create-a

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare-a
#| eval: false
waldo::compare(solar, solution_a, tolerance = 1e-4, ignore_attr = c("class", "groups"))
```


### Part 2 - Solar panel angles

Next, let's examine the effect of angling a solar panel towards the sun. NASA's data contains three measures of solar illuminance:

1. `direct_lux` - the amount of illumination received directly from the solar disk by a panel optimally _angled towards the sun_
2. `diffuse_lux` - the amount of illumination received indirectly (from sources other than the solar disk) by a panel laid _flat on the ground_
3. `global_lux` - the total amount of illumination, direct + diffuse, received by a panel laid _flat on the ground_

Run the code below to see a table that contains two new variables:

1. `flat_lux` - the amount of illumination received directly from the solar disk by a panel laid _flat on the ground_
2. `gain` - the difference between `direct_lux` and `flat_lux`

```{r}
#| label: recreate-this-b
#| message: false
solution_b <- readr::read_csv("data/milestone03b.csv")
solution_b
```

Your task is to use what you've learned about the dplyr package to transform `solar` into this table.

Work in the code chunk below. Save the result as `gains`. We will use `gains` in Part 3.

*Hint*: Pay attention to the position of these new variables in the solution table -- they immediately follow the `global_lux` and `direct_lux` columns.

```{r}
#| label: recreation-create-b

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare-b
#| eval: false
waldo::compare(gains, solution_b, tolerance = 1e-4, ignore_attr = c("class", "groups"))
```

### Part 3 - Visualize it

The amount of gain to be had by angling a solar panel varies from day to day. What does the distribution of gains look like?

Using the File Explorer, navigate to the `images/` directory, then click on `milestone03a.png` to open an image.

Use the code chunk below to create this plot. Begin with the `gains` table.

```{r}
#| label: recreation-plot-a

```

#### Clearness

Perhaps angling a solar panel creates a larger gain on days when the sky is clear, and a smaller gain on days when it is not. We can test this visually with the `clearness` variable. `clearness` measures how clear the sky is, with 1 being completely clear and 0 being opaque.

Using the File Explorer, navigate to the `images/` directory, then click on `milestone03b.png` to open an image.

How do you interpret the results?

Use the code chunk below to create this plot. Begin with the `gains` table.

```{r}
#| label: recreation-plot-b

```

## Extension

Using the code chunk below, investigate a research question about this data, using the additional data wrangling skills you learned this week. Some ideas:

1. How does the gain from angling a solar panel vary by month (or some other unit of time of interest)?
2. Based on your new `gain` variable, split the data into groups of your choosing (e.g. low/medium/high) and explore patterns based on these groups.
3. Is there a relationship between efficiency gain and the degree of ultraviolet radiation exposure?
4. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. Create at least one new variable in your data set (using `mutate()` and/or `summarize()`), or use other dplyr functions to alter your data
3. Use your updated data set to create at least one graph and/or table


```{r}
#| label: extension

```
