---
title: "Work with data types"
format: html
---

# Instructions

This week, you worked with three new data types in R -- factors, strings, and date-times -- using the forcats, stringr, and lubridate packages, respectively.

In this milestone, you'll use the date-time information in our data to plot the average precipitation by season.

To learn more about the variables contained in the data set(s), refer to the data dictionaries: https://rsacdn.link/milestones/internal/v2/tidyverse/solar-power/assets/illuminance_dictionary.html and https://rsacdn.link/milestones/internal/v2/tidyverse/solar-power/assets/precip_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false
```

## Data types

In this milestone, you'll use lubridate to manipulate the date information in our data and create a season variable.

## Recreation

### Part 1 - Import

Before you begin, you will need to import your data sets. Use the code chunk below to read two data sets from the `data/` directory:

1. `illuminance.csv`, save it as `illuminance`
2. `precip.csv`, save it as `precip`

```{r}
#| label: recreation-import

```

Next, run the code chunk below to recreate `solar` from `illuminance`. This version of `solar` also contains the variables you made in previous milestones, as well as the data you joined from `precip`.

```{r}
#| label: data-prep
solar <-
  illuminance |>
  rename(
    global_lux = global_illuminance,
    diffuse_lux = diffuse_illuminance,
    direct_lux = direct_illuminance,
    pressure = ps,
    temp = t2m,
    temp_min = t2m_min,
    temp_max = t2m_max,
    dew_point = t2mdew,
    snow = snodp,
    wind_speed = ws2m,
    wind_direction = wd2m,
    uv_index = allsky_sfc_uv_index,
    albedo = allsky_srf_alb,
    optical_depth_55 = aod_55,
    optical_depth_84 = aod_84,
    clearness = clrsky_kt
  ) |>
  mutate(
    flat_lux = global_lux - diffuse_lux,
    gain = direct_lux - flat_lux,
    .after = diffuse_lux
  ) |>
  left_join(precip, by = c("lat", "lon", "year", "mm", "dd", "doy")) |>
  rename(
    precip = prectotcorr,
    humidity_specific = qv2m,
    humidity_relative = rh2m,
    cloud_depth = cloud_od,
    water = pw
  ) |>
  select(-c(lon, lat))
```

### Part 2 - Dates

Run the code below to see a table.

```{r}
#| label: recreate-this-1
#| message: false
solution <- readr::read_csv("data/milestone05.csv")
solution
```

Your task is to use what you've learned about the lubridate package to transform `solar` into this table. You will need to:

1. Combine the year, month, and day information into a new variable named `date` using the `make_date()` function
2. Save the month of each date, as a label, in a new variable named `month`
3. Save the quarter of each date in a new variable named `quarter`
4. Recode the values of `quarter` to make a new variable named `season`. Recode the values such that:

    +  Quarter 1 = "Winter"
    +  Quarter 2 = "Spring"
    +  Quarter 3 = "Summer"
    +  Quarter 4 = "Fall"

5. Add your new variables before the `doy` column of `solar`

Work in the code chunk below. Save the result as `solar_dates`; we will use it in Part 3.

```{r}
#| label: recreation-seasons

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(solar_dates, solution, tolerance = 1e-4, ignore_attr = c("class", "groups"))
```

## Part 3 - The role of factor orders in visualizations

Now that you have a `season` variable, you can compute and visualize summaries for each season. Run the chunk below to view a summary table.

```{r}
#| label: summary
seasons_summary <-
  solar_dates |>
  group_by(season) |>
  summarize(
    avg_precip = mean(precip, na.rm = TRUE)
  )

seasons_summary
```

After you have run the chunk above to create `seasons_summary`, run the chunk below to visualize the data of `seasons_summary`.

```{r}
#| label: summary-plot
#| eval: false
seasons_summary |>
  ggplot(mapping = aes(x = season, y = avg_precip)) +
  geom_col() +
  labs(title = "Average Precipitation by Season")
```

This is a nice plot, but the order of seasons on the x axis is arbitrary (alphabetical). Let's give the seasons a more principled order before we publish this plot.

#### Visualization

Using the File Explorer, navigate to the `images/` directory, then click on `milestone05.png` to open an image.

You should see a plot in which the seasons are now ordered from most to least rainiest. Your task is to recreate this plot.

Use the chunk below to recreate this plot. Hint: turn the `season` variable into a factor with a certain order.

```{r}
#| label: recreation-visualize-1

```

## Extension

Using the code chunk below, investigate a research question about this data, using the skills you have learned this week and in previous weeks of Academy. Some ideas:

1. Use data visualization to investigate if there is a seasonal pattern in other variables we've computed in previous milestones, e.g. efficiency gain
2. Explore potential seasonal patterns in solar or meteorological variables you have not yet investigated.
3. [any other research question of interest]


Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. If your data contains ...
  ... date-time data, use a lubridate function to correctly format your date-time variables
  ... string data, use a stringr function to manipulate your string variables
  ... categorical data, use a forcats function to modify the factor levels of your categorical variables (reorder, recode, etc.)
3. Using a table or graph, present a meaningful subset of your modified data

```{r}
#| label: extension

```
