---
title: "Tidy and join tables"
format: html
---

# Instructions

This week you learned about tidying and joining data sets with the tidyr (pronounced tidy-R) and dplyr packages.

Joining data sets can be deceptively tricky. It is important to keep in mind that prepping and inspecting the tables you're working with *before* you do a join is often more important (and takes more time) than doing the join itself.

You'll use some of these skills in the milestone to the right.

To learn more about the variables contained in the data set(s), refer to the data dictionaries: https://rsacdn.link/milestones/internal/v2/tidyverse/solar-power/assets/illuminance_dictionary.html and https://rsacdn.link/milestones/internal/v2/tidyverse/solar-power/assets/precip_dictionary.html

Remember that all tidyverse packages have already been installed for you but you will need to *load* the ones you want to use.

# Milestone

```{r}
#| label: setup
#| include: false

# Load your packages here

```

## Tidy and join

In this milestone, you'll check how specifically cloudy conditions affect direct and diffuse illuminance. To do this, you'll need to use the data in `precip.csv` alongside the data you've been using in `illuminance.csv`. Both data sets are stored in your `data/` directory.

You will also use pivoting to directly compare `flat_lux` to `direct_lux` in a plot.

## Recreation

### Part 1 - Import

Before you begin, you will need to import your data sets. Use the code chunk below to read two data sets from the `data/` directory:

1. `illuminance.csv`, save it as `illuminance`
2. `precip.csv`, save it as `precip`

```{r}
#| label: recreation-import

```

Next, run the code chunk below to recreate `solar` from `illuminance`. Notice that this version of `solar` also contains the `flat_lux` and `gain` variables you made in the previous milestone.

```{r}
#| label: solar
solar <-
  illuminance |>
  rename(
    global_lux = global_illuminance,
    diffuse_lux = diffuse_illuminance,
    direct_lux = direct_illuminance,
    pressure = ps,
    temp = t2m,
    temp_min = t2m_min,
    temp_max = t2m_max,
    dew_point = t2mdew,
    snow = snodp,
    wind_speed = ws2m,
    wind_direction = wd2m,
    uv_index = allsky_sfc_uv_index,
    albedo = allsky_srf_alb,
    optical_depth_55 = aod_55,
    optical_depth_84 = aod_84,
    clearness = clrsky_kt
  ) |>
  mutate(
    flat_lux = global_lux - diffuse_lux,
    gain = direct_lux - flat_lux,
    .after = diffuse_lux
  )
```

### Part 2 - Angles

In the last milestone, we computed the daily performance gain in direct illuminance that a solar panel angled optimally towards the sun has over a panel laid flat on the ground.

Using the File Explorer, navigate to the `images/` directory, then click on `milestone04.png` to open an image.

Use the code chunk below to create this plot. Begin with the `solar` table.

*Hint*: You will need to summarize the `solar` table and pivot it into a longer format before plotting.

```{r}
#| label: recreation-plot-a

```

### Part 3 - Join tables

Run the code below to see a table.

```{r}
#| label: recreate-this
#| message: false
solution <- readr::read_csv("data/milestone04.csv")
solution
```

Your task is to use `precip` and what you've learned about the dplyr package to transform `solar` into this table. You may need to do some pre-join checks:

1. Which variables can you join on?
2. Do the data sets have any mismatched observations?
3. Perform any necessary data cleaning.
4. Check again as needed!

*Hints*:
* Your table should no longer have the `lat` and `lon` variables
* Learn more about the variables in `precip` in the data dictionary: https://rsacdn.link/milestones/internal/v2/tidyverse/solar-power/assets/precip_dictionary.html
* You will need to rename some columns in `precip` -- replace the column names on the right with the new names on the left:

| New name          | Old name    |
|-------------------|-------------|
| precip            | prectotcorr |
| humidity_specific | qv2m        |
| humidity_relative | rh2m        |
| cloud_depth       | cloud_od    |
| water             | pw          |

Work in the code chunk below. Save the result as `solar_sky`; we will use it in Part 4.

```{r}
#| label: recreation-join

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(solar_sky, solution, tolerance = 1e-4, ignore_attr = c("class", "groups"))
```

### Part 4 - Visualize

Let's explore our results visually. What do you think the code chunks below will do? Run the chunks to see if you were right.

```{r}
#| label: plot
#| eval: false
solar_sky |>
  ggplot(aes(x = cloud_amt, y = direct_lux)) +
  geom_point()
```

```{r}
#| label: plot2
#| eval: false
solar_sky |>
  ggplot(aes(x = cloud_amt, y = diffuse_lux)) +
  geom_point()
```

How do you interpret the results?

## Extension

Using the code chunk below, investigate a research question about this data, using the skills you have learned this week and in previous weeks of Academy. Some ideas:

1. Create a visualization using the joined data from the milestone recreation. Use this visualization as the starting point for your own further investigation of the data. Use at least one data transformation step and one graph to start addressing your questions.
2. In the recreation you explored the relationship between cloudiness and illuminance. Do you see a relationship between illuminance and other meteorological variables (e.g. precipitation, humidity, etc.)?
3. Consider finding and joining in other publicly available meteorological or environmental data, and ask a question of the new data set.
4. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. Complete a data wrangling task relevant to your data (tidying and/or reshaping/pivoting your data); if appropriate, join your data with a second data set (filtering or mutating join)
3. Show summary tables of your data throughout these transformations


```{r}
#| label: extension

```
