---
title: "Summarize data"
format: html
---

# Instructions

This week, you learned several new functions from dplyr (pronounced DEE-ply-er), a commonly-used package from the tidyverse for wrangling data. You also learned some new ways to enhance your plots.

You'll use these skills to quantify the seasonal trend that we saw in the last milestone, and double check that every year follows the trend.

To learn more about the variables contained in the data set, refer to the data dictionary: https://rsacdn.link/milestones/internal/v2/tidyverse/solar-power/assets/illuminance_dictionary.html

# Milestone

```{r}
#| label: setup
#| include: false

# Load your packages here

```

## Summarize data

In this milestone you'll create, visualize, and compare summary statistics that estimate the solar potential for each month of the year.

## Recreation

### Part 1 - Import

Before you begin, you will need to import your data set. Use the code chunk below to read the data from the data file `illuminance.csv`, which is stored in the `data/` folder in your working directory. Be sure to save the data to an object named `illuminance`.

```{r}
#| label: recreation-import

```

### Part 2 - Seasonal trends by year

In the last milestone, you plotted the seasonal trend for 2013. Now let's plot the seasonal trend for _every_ year in the data set.

Using the File Explorer, navigate to the `images/` directory, then click on `milestone02a.png` to open an image.

Use the code chunk below to create this plot. Begin with the `illuminance` table.

```{r}
#| label: recreation-plot-a

```

Now, let's overlay the trends on top of one another, to check whether any years defied the trend.

Using the File Explorer, navigate to the `images/` directory, then click on `milestone02b.png` to open an image.

Use the code chunk below to create this plot. Begin with the `illuminance` table.

```{r}
#| label: recreation-plot-b

```

### Part 3 - Summarize the data by month

Run the code below to see a table.

```{r}
#| label: recreate-this
#| message: false
solution <- readr::read_csv("data/milestone02.csv")
solution
```

Your task is to use what you've learned about the dplyr package to transform `illuminance` into this table. You will need to:

1. Group the data
2. Summarize `global_illuminance` for each group

Work in the code chunk below. Save the result as `monthly_averages`. We will use the result in Part 4.

```{r}
#| label: recreation-summary

```

Run the following code chunk to test whether you have the same answer as the solution:

```{r}
#| label: compare
#| eval: false
waldo::compare(monthly_averages, solution, tolerance = 1e-4, ignore_attr = c("class", "groups"))
```

### Part 4 - Visualize

Let's display our results visually.

Using the File Explorer, navigate to the `images/` directory, then click on `milestone02c.png` to open an image.

Use the code chunk below to create this plot. Begin with the `monthly_averages` table.

```{r}
#| label: recreation-plot-c

```

## Extension

Using the code chunk below, investigate a research question about this data, using the data wrangling skills you learned this week. Some ideas:

1. In addition to monthly averages of global illuminance, calculate and visualize daily and yearly averages. How much variation do you see in illuminance within each of these levels of temporal granularity?
2. What temporal patterns do you see in other variables, e.g. surface pressure, snow depth, surface temperature, etc.? Do you notice different temporal variations across these different variables? Does this track with your intuition?
3. [any other research question of interest]

Alternately, working with a data set of your own, complete the following:

1. Read in your data
2. Use `group_by()` and `summarize()` to produce simple a table from some subset of your data
3. Create a related graph of this data subset using ggplot

```{r}
#| label: extension

```
